```{r}
library(readr)
library(conText)
library(dplyr)
library(tidytext)
library(quanteda)
library(word2vec)

corpus <- read_csv("~/Desktop/UNGDC/data/processed/cleaned.csv") %>%
  select(-"...1")
glove <- readRDS("~/Desktop/UNGDC/data/embeddings/glove.rds")
meta <- read_csv("~/Desktop/UNGDC/data/processed/meta.csv") 

corpus <- corpus(meta)
toks <- tokens(corpus, 
               remove_punct=T, 
               remove_symbols=T, 
               remove_numbers=T, 
               remove_separators=T)

toks_nostop <- tokens_select(toks, 
                             pattern = stopwords("en"), 
                             selection = "remove", min_nchar=3)

feats <- dfm(toks_nostop, 
             tolower=T, 
             verbose = FALSE) %>% 
  dfm_trim(min_termfreq = 5) %>% featnames()

toks_nostop_feats <- tokens_select(toks_nostop, feats, padding = TRUE)


sov_tokens <- tokens_context(x = toks_nostop_feats, pattern = "soverei*", window = 10L)
sov_tokens <- sov_tokens[!is.na(docvars(sov_tokens)$dd_regime), ]

head(docvars(sov_tokens),3)

sov_dfm <- dfm(sov_tokens)

# build a document-embedding-matrix
sov_dem <- dem(x = sov_dfm, 
               pre_trained = glove, 
               transform = TRUE, 
               transform_matrix = cr_transform, 
               verbose = TRUE)


sov_wv <- matrix(colMeans(sov_dem), ncol = ncol(sov_dem)) %>% 
  `rownames<-`("sovereignty")

sov_wv_regime <- dem_group(sov_dem,
                           groups = sov_dem@docvars$dd_regime)

nns(sov_wv, pre_trained = glove, N = 25, candidates = sov_wv_regime@features, stem =FALSE, as_list = FALSE)
```
#Nearest neighbors
Numeric regime category, following Cheibub, Ghandi and Vreeland (2010) (Parliamentary democracies = 0, Mixed democracies (with weak presidents) = 1, Presidential democracies = 2, Civilian autocracies = 3, Military dictatorships = 4, and Royal dictatorships = 5)
```{r}
feats <- featnames(dfm(sov_tokens))

set.seed(2024)

sov_tokens <- sov_tokens[!is.na(docvars(sov_tokens)$dd_regime), ]


sov_regime_nns <- get_nns(x = sov_tokens, N = 20,
                          groups = (docvars(sov_tokens, 
                                          'dd_regime')),
                          candidates = feats,
                          pre_trained = glove,
                          transform = TRUE,
                          transform_matrix = cr_transform,
                          bootstrap = TRUE,
                          num_bootstraps = 100,
                          confidence_level = 0.95,
                          as_list = TRUE
                          )
sov_regime_nns[["0"]]
sov_regime_nns[["1"]] 
sov_regime_nns[["5"]] #full dic


```
#cosine similarity
```{r}
# compute the cosine similarity between each group's embedding and a specific set of features
set.seed(2024)
get_cos_sim(x = sov_tokens,
            groups = docvars(sov_tokens, 'dd_regime'),
            features = c("right", "security", "border", "independence"),
            pre_trained = glove,
            transform = TRUE,
            transform_matrix = cr_transform,
            bootstrap = TRUE,
            num_bootstraps = 100,
            as_list = FALSE)
```
#nns_ratio
```{r}
set.seed(2024)
sov_nns_ratio <- get_nns_ratio(x = sov_tokens, 
              N = 20,
              groups = docvars(sov_tokens, 'democracy'),
              numerator = "1",
              candidates = feats,
              pre_trained = glove,
              transform = TRUE,
              transform_matrix = cr_transform,
              bootstrap = TRUE,
              num_bootstraps = 100,
              permute = TRUE,
              num_permutations = 100,
              verbose = FALSE)

```

#nearest contexts
computing cosine similarities between group embeddings and individual contexts

```{r}
set.seed(2024)
sov_ncs <- ncs(x = sov_wv_regime,
               contexts_dem = sov_dem, 
               contexts = sov_tokens, 
               N=10,
               as_list = TRUE)
```

# stemming option turned on
```{r}
# find nearest neighbors by party using stemming
sov_nns_stem <- nns(sov_wv_regime, pre_trained = glove, N = 5, candidates = feats, stem = TRUE, as_list = TRUE)
```

#regression
```{r}
set.seed(2024)
model1 <- conText(formula =  . ~ dd_regime,
                  data = toks_nostop_feats,
                  pre_trained = glove,
                  transform = TRUE, transform_matrix = cr_transform,
                  bootstrap = TRUE,
                   confidence_level = 0.95,
                  permute = TRUE, num_permutations = 100,
                  window = 6, case_insensitive = TRUE,
                  verbose = FALSE)
```
